Location and Scale-Invariant Power Transformations for Transforming Data to Normality

### 2023-11-15 manuscript/manuscript.Rmd
Work on manuscript.

# Set general theme.
plot_theme <- ggplot2::theme_light(base_size=base_size)
plot_theme$strip.background <- ggplot2::element_blank()
plot_theme$strip.text$colour <- NULL
plot_theme$plot.margin <- grid::unit(c(2, 2, 2, 2), "points")

@@

```{r decreased-normality, echo=FALSE, fig.cap=cap, fig.height=4.0, warning=FALSE}
cap <- paste0(
  "Power transformations may reduce normality depending on location. ",
  "Variable $x$ consists of $10000$ values randomly drawn from a normal distribution. ",
  "For the top row, that distribution is parameterised as $\\mathcal{N}(\\mu, 1)$, ",
  "whereas for the bottom row, the distribution is $\\mathcal{N}(0, \\sigma)$.",
  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
  "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
  "and breaks down due to numerical issues for $\\mu > 10^3$ (both Box-Cox and Yeo-Johnson) ",
  "and $\\sigma > 10^3 (Yeo-Johnson) only."
)

@@

## Shift and scale invariant power transformation

Box-Cox and Yeo-Johnson transformations are modified by introducing shift parameter $x_0$ and scale parameter $s$ into equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}. The shift and scale invariant Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is then:

@@

The result is shown in Figure \ref{fig:shifted-distributions}. For each distribution, transformation parameter $\hat{\lambda}$ varies with $d_{\text{shift}}$ and $d_{\text{scale}}$ when estimated using the original definitions. In contrast, estimation of $\lambda$ using invariant power transformations was invariant to $d_{\text{shift}}$ and $d_{\text{scale}}$.

>>>>

# Set general theme.
plot_theme <- ggplot2::theme_light(base_size=base_size)
plot_theme$plot.title$size <- ggplot2::rel(1.0)
plot_theme$strip.background <- ggplot2::element_blank()
plot_theme$strip.text$colour <- NULL
plot_theme$plot.margin <- grid::unit(c(2, 2, 2, 2), "points")

@@

```{r decreased-normality, echo=FALSE, fig.cap=cap, fig.height=4.0, warning=FALSE}
cap <- paste0(
  "Power transformations may reduce normality depending on location. ",
  "Variable $x$ consists of $10000$ values randomly drawn from a normal distribution. ",
  "For the left-hand panels, that distribution is parameterised as $\\mathcal{N}(\\mu, 1)$, ",
  "whereas for the right-hand row, the distribution is $\\mathcal{N}(0, \\sigma)$.",
  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
  "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
  "and breaks down due to numerical issues for $\\mu > 5 \\cdot 10^3$ (both Box-Cox and Yeo-Johnson).",
  "Scale did not affect Box-Cox and Yeo-Johnson transformations of variables drawn from normal distributions."
)

@@

## Location and scale invariant power transformation

Box-Cox and Yeo-Johnson transformations are modified by introducing shift parameter $x_0$ and scale parameter $s$ into equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}. The location and scale invariant Box-Cox transformation of a feature value $x_i$ of feature $\mathbf{X}$ under transformation parameter $\lambda$, shift parameter $x_0$ and scale parameter $s$ is then:

@@

The result is shown in Figure \ref{fig:shifted-distributions}. For each distribution, transformation parameter $\hat{\lambda}$ varies with $d_{\text{shift}}$ and $d_{\text{scale}}$ when estimated using the original definitions. In contrast, estimation of $\lambda$ using invariant power transformations was invariant to $d_{\text{shift}}$ and $d_{\text{scale}}$. While the effect of location on the estimated value of $\lambda$ using conventional power transformations is noticeable, scale also affects $\lambda$ for features that are not normally distributed.
The figure also shows that estimated $\lambda$ for Box-Cox power transformations in left-skewed distributions diverges from that found for invariant Box-Cox power transformations.
